Parsytec reserves the right to make changes in specification at any time and 
without notice. The information furnished by Parsytec in this publication is
believed to be accurate, however no responsibility is assumed for its use, nor
for any infringement of patents or rights of third parties resulting from its 
use. No license is granted under any patents, trademarks or other rights of
Parsytec.

\vfill

This manual is Copyright \copyright\ 1992 by Parsytec GmbH \\
The Parsytec File System is Copyright \copyright\ 1988 - 1992
by Parsytec GmbH.

Authors:\\
\begin{tabular}{ll}
  Adruni Ishan        & basic Version (design and implementation)      \\
  Matthias Clauss     & basic Version (design)                         \\   
  Hans-Joachim Ermen  & checker                                        \\ 
  Christian Fleischer & MSC device driver, {\tt ptar} utility          \\
  G\"unter Lauven     & multivolume extensions and checker integration \\
  Oliver Imbusch      & v2.1 bug fixes and minor extensions            \\
\end{tabular}

This documentation was written by Oliver Imbusch, \\
contributions by A Ishan (Version 1.1 Handbook), H-J Ermen (Checker
Documentation), N Garnett (Technical Report No 20 of Perihelion Software,
``Device Configuration''), G Lauven (``Short description of PFS v2.0''),
and C Fleischer (``MSC Device Driver Documentation'', ``The {\tt ptar} 
utility'').

All Rights reserved. This document may not, in whole or in part, be copied, 
photocopied, reproduced or reduced to any electronic medium or machine
readable form without prior consent, in writing from\\

Parsytec GmbH\\
Juelicher Strasse 338\\
D-5100 Aachen\\
Germany.\\

HELIOS\trademark\ is a trademark of Perihelion Software Ltd.\\
SUN Workstation \registered\ is a registered trademark of Sun Microsystems,
Inc.\\
Occam \registered\ is a registered trademark of the INMOS Group of Companies.

The use of registered names, trademarks, etc. in this publication does not
imply, even in the absence of a specific statement, that such names are exempt
from relevant protective laws and regulations and therefore free for general use.
